Evaluation of probabilistic photometric redshift estimation approaches for \boldblue{ \textsc{The Rubin Observatory Legacy Survey of Space and Time} (LSST)}

### 2020-09-03 paper/tex/analysis.tex
redefine macros, fix a few spacing issues

The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, \mathrm{data})$.
To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
The conditional density estimation (CDE) loss is an analogue to the familiar root-mean-square-error used in conventional regression, defined as
\begin{equation}
  \label{eq:cde-loss}
  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} \mathrm{d}z \mathrm{d}P(\x) ,
\end{equation}
where $f(z \vert \x)$ is the true \pzpdf\ that we do not have and $\widehat{f}(z \vert \x)$ is an estimate thereof, in terms of the photometry $\x$.
(See Section~\ref{sec:flexzboost} for a review of the notation.)
We estimate the CDE loss via
\begin{equation}
  \label{eq:estimated-cde-loss}
  \hat{L}(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X, Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
\end{equation}
where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$.
\sout{We may estimate these expectations empirically on the test or validation data} \citep[Eq.~7 in][]{Izbicki:17b} \sout{without knowledge of the true densities.}
\boldblue{One of the most powerful features of the CDE loss is the ability to estimate the loss function up to a constant even in the absence of knowledge of the true underlying distribution (see Eq.~7 and accompanying discussion in \citet{Izbicki:17} for more details).  This feature enables a quantitative comparison of our estimation methods with our current dataset where we lack access to the true $f(z \vert \x)$.}

>>>>

The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, \mathrm{data})$.
To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
The conditional density estimation (CDE) loss is an analogue to the familiar root-mean-square-error used in conventional regression, defined as
\begin{equation}
  \label{eq:cde-loss}
  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} \mathrm{d}z \mathrm{d}P(\x) ,
\end{equation}
where $f(z \vert \x)$ is the true \pzpdf\ that we do not have and $\widehat{f}(z \vert \x)$ is an estimate thereof, in terms of the photometry $\x$.
(See Section~\ref{sec:flexzboost} for a review of the notation.)
We estimate the CDE loss via
\begin{equation}
  \label{eq:estimated-cde-loss}
  \hat{L}(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X, Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
\end{equation}
where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$\xsout{We may estimate these expectations empirically on the test or validation data} \citep[Eq.~7 in][]{Izbicki:17b}. \xsout{without knowledge of the true densities.}
\boldblue{One of the most powerful features of the CDE loss is the ability to estimate the loss function up to a constant even in the absence of knowledge of the true underlying distribution (see Eq.~7 and accompanying discussion in \citet{Izbicki:17} for more details).  This feature enables a quantitative comparison of our estimation methods with our current dataset where we lack access to the true $f(z \vert \x)$.}

### 2020-09-03 paper/tex/data.tex
redefine macros, fix a few spacing issues

The full dataset thus covers $400$ square degrees and contains $238$ million galaxies of redshift $0 < z \leq 8.7$ down to $r = 29$.  \boldblue{However, a problem with the k-correction calculation led to incorrect colors at $z>2$, which necessitated limiting the redshift range to $0 < z \leq 2.0$.}
\sout{Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.}
To obtain a catalogue matching the \lsst\ Gold Sample, we imposed a cut of $i < 25.3$, which gives a signal-to-noise ratio $\gtrsim 30$ for most galaxies.
In order for statistical errors to be subdominant to the systematic errors we aim to probe, we further reduced the sample size to $<10^{7}$ galaxies by isolating $\sim 16.8$ square degrees selected from five separate spatial regions of the simulation.
We refer to this final set of galaxies as DC1, for the first \lsstdesc\ Data Challenge.

>>>>

The full dataset thus covers $400$ square degrees and contains $238$ million galaxies of redshift $0 < z \leq 8.7$ down to $r = 29$.  \boldblue{However, a problem with the k-correction calculation led to incorrect colors at $z>2$, which necessitated limiting the redshift range to $0 < z \leq 2.0$.}
\xsout{Systematic inconsistencies with galaxy colors at $z > 2$ were observed, so the catalogue was limited to $0 < z \leq 2.0$.}
To obtain a catalogue matching the \lsst\ Gold Sample, we imposed a cut of $i < 25.3$, which gives a signal-to-noise ratio $\gtrsim 30$ for most galaxies.
In order for statistical errors to be subdominant to the systematic errors we aim to probe, we further reduced the sample size to $<10^{7}$ galaxies by isolating $\sim 16.8$ square degrees selected from five separate spatial regions of the simulation.
We refer to this final set of galaxies as DC1, for the first \lsstdesc\ Data Challenge.

### 2020-09-03 paper/tex/methods.tex
redefine macros, fix a few spacing issues

\sout{Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.}

@@

We test three publicly available and commonly used template-based codes that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a \chisq\ merit function using the observed and model fluxes:
\begin{equation} \label{eq_temp_chi}
\chi^{2}(z,T,A) = \sum_{i}^{N_{\mathrm{filt}}}\left(\frac{F^{i}_{\mathrm{obs}} - A \, F^{i}_{\mathrm{pred}}(T,z)}{\sigma^{i}_{\mathrm{obs}}}\right)^2
\end{equation}
\noindent where $A$ is a normalization factor, $F^i_{\mathrm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\mathrm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\mathrm{obs}}$ is the observed flux error, and $N_{\mathrm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.  \boldblue{All three template-based codes replace per-band non-detections with an estimate of the 1$\sigma$ magnitude limit as estimated from the photometric error model.}
\sout{The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which are the \lsst\ $ugrizy$ filters in this case.}
Specific implementation details of each code, e.~g.~prior form and implementation, are described below.

@@

In this study, we use \annz-v.2.0.4.  \boldblue{Several combinations of the ANN, BDT, and kNN machine learning algorithms were tested during the training/validation stage.  For this particular dataset, the optimal and most stable results came from a setup consisting solely of ANNs.  We present results for this case and do not include results from a mixture of multiple algorithms. }
\sout{to output only the result of the ANN algorithm.}
\Pzpdf s were produced by running an ensemble of 5 ANNs with a $6:12:12:1$ architecture corresponding to the 6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output of redshift.
Each of the five ANNs was trained with different random seeds for the initialization of input parameters, reserving half of the training set for validation to prevent overfitting.
\boldblue{Galaxies not detected in the $u$-band} \sout{Undetected galaxies} were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude \boldblue{of the test set} in that band. \sout{within the entire test set.}

@@

The colour-matched nearest-neighbours photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c_{\mathrm{j,train}} - c_{\mathrm{j,test}})^{2} / \delta c_{\mathrm{j,test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colours $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
\sout{Undetected bands} \boldblue{Bands in which the galaxy is not detected} are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.

>>>>

\xsout{Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.}

@@

We test three publicly available and commonly used template-based codes that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a \chisq\ merit function using the observed and model fluxes:
\begin{equation} \label{eq_temp_chi}
\chi^{2}(z,T,A) = \sum_{i}^{N_{\mathrm{filt}}}\left(\frac{F^{i}_{\mathrm{obs}} - A \, F^{i}_{\mathrm{pred}}(T,z)}{\sigma^{i}_{\mathrm{obs}}}\right)^2
\end{equation}
\noindent where $A$ is a normalization factor, $F^i_{\mathrm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$, $F^i_{\mathrm{obs}}$ is the observed flux in a given band $i$, $\sigma^i_{\mathrm{obs}}$ is the observed flux error, and $N_{\mathrm filt}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.  \boldblue{All three template-based codes replace per-band non-detections with an estimate of the 1$\sigma$ magnitude limit as estimated from the photometric error model.}
\xsout{The likelihood is a sum of observed flux error $\sigma_{b}^{\mathrm{obs}}$-weighted squared differences between the observed flux $F^{\mathrm{obs}}_{b}$ and the normalized predicted flux $F^{\mathrm{mod}}_{b}(T, z)$ in $N_{\mathrm{filt}}$ photometric filters $b$, which are the \lsst\ $ugrizy$ filters in this case.}
Specific implementation details of each code, e.~g.~prior form and implementation, are described below.

@@

In this study, we use \annz-v.2.0.4.  \boldblue{Several combinations of the ANN, BDT, and kNN machine learning algorithms were tested during the training/validation stage.  For this particular dataset, the optimal and most stable results came from a setup consisting solely of ANNs.  We present results for this case and do not include results from a mixture of multiple algorithms. }
\xsout{to output only the result of the ANN algorithm.}
\Pzpdf s were produced by running an ensemble of 5 ANNs with a $6:12:12:1$ architecture corresponding to the 6 $ugrizy$ inputs, 2 hidden layers with 12 nodes each, and 1 output of redshift.
Each of the five ANNs was trained with different random seeds for the initialization of input parameters, reserving half of the training set for validation to prevent overfitting.
\boldblue{Galaxies not detected in the $u$-band} \xsout{Undetected galaxies}were excluded from the training set, and per-band non-detections in the test set were replaced with the mean magnitude \boldblue{of the test set} in that band. \xsout{within the entire test set.}

@@

The colour-matched nearest-neighbours photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c_{\mathrm{j,train}} - c_{\mathrm{j,test}})^{2} / \delta c_{\mathrm{j,test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colours $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
\xsout{Undetected bands}\boldblue{Bands in which the galaxy is not detected} are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.
